---
title: "TwinMom 16S Analysis"
author: "Rachel Rodgers"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r global-options, warning=FALSE, message=FALSE, echo=FALSE}
# TwinMom 16S script 4

knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      cache = T)
```

```{r load-libraries}
library("phyloseq")
library("plotly")
library("microbiome")
library("decontam")
library("ggpubr")
library("ggrepel")
library("vegan")
library("pheatmap")
library("reshape2")
library("scales")
library("data.table")
library("qpcR")
library("RColorBrewer")
library("gridExtra")
library("DESeq2")
library("tidyverse")

source("./shared_R_scripts/Helper_Functions.R")
source("./TwinMom16S_Shared_Functions.R")
```

# Intro

In this scipt we analyze the diversity of 16S stool samples.

```{r load-physeqStool}
physeqStool <- readRDS("../../data/physeqObjects/physeqStool.RDS")
```

# Alpha Diversity Plots

```{r ecoIdxDF, fig.width=9, fig.height=6}
ecoIdxDF <- global(x = physeqStool, 
                   index = c("observed", "diversities_shannon"))
metadataStool <- data.frame(sample_data(physeqStool))
ecoIdxDF <- merge(metadataStool, ecoIdxDF, by = "row.names")

AlphaBoxPlot <- function(df, 
                        xVar, 
                        yVar, 
                        label_y = NULL, 
                        label_x = NULL, 
                        statMethod = NULL) {

  ggplot(data = df, aes_string(x = xVar, y = yVar)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.2) +
    ylab(label_y) +
    xlab(label_x) +
    theme_pubr() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_text(size = 20),
          axis.text.x = element_text(size = 18),
          axis.text.y = element_text(size = 18)) +
      stat_compare_means(method = statMethod, label.x.npc = 0.5, size = 5)
}

# Rename levels of sample type
levels(ecoIdxDF$SampleType)
ecoIdxDF$SampleType <- factor(ecoIdxDF$SampleType,
                              labels = c("Infants", "Mothers"))

richnessPlot <- AlphaBoxPlot(df = ecoIdxDF,
                            xVar = "SampleType",
                            yVar = "observed",
                            label_y = "Richness",
                            statMethod = "wilcox.test")

diversityPlot <- AlphaBoxPlot(df = ecoIdxDF,
                             xVar = "SampleType",
                             yVar = "diversities_shannon",
                             label_y = "Shannon Diversity",
                             statMethod = "wilcox.test")

richnessPlot
diversityPlot
```

# Community Composition Plots

```{r familyCompPlot, fig.width=11, fig.height=8.5}
# Calculate abundance data frame at Family rank
familyAbdDF <- MakeAbundanceDF(physeq = physeqStool,
                               taxRank = "Family",
                               abundanceFilter = 0)

# Order infant and mother samples by Tarr ID
metadataStool <- metadataStool %>%
  rownames_to_column(var = "Sample") %>%
  arrange(InfantMother, Tarr_Subject_ID)

sampleLevels <- as.character(metadataStool$Sample)

sampleLabels <- metadataStool %>%
  dplyr::select(Sample, Tarr_Subject_ID, InfantMother)

sampleLabels$InfantMother <- factor(sampleLabels$InfantMother,
                                    labels = c("I", "M"))

sampleLabels <- sampleLabels %>%
  mutate("ID" = paste0(InfantMother, Tarr_Subject_ID))

sampleLabelsVec <- sampleLabels$ID
names(sampleLabelsVec) <- sampleLabels$Sample

# Order the samples in familyAbdDF by the sampleLevels
familyAbdDF$Sample <- factor(familyAbdDF$Sample, levels = sampleLevels)

familyCompPlot <- ggplot(familyAbdDF,
                         aes_string(x = "Sample", y = "Abundance",
                                    fill = "Family")) +
  geom_bar(stat = "identity", width = 1) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(labels = sampleLabelsVec)
familyCompPlot
```

# PCoA of Weighted UniFrac Distances

```{r calculate-ord-objects}
set.seed(787)
unweightedOrd <- ordinate(physeqStool, method = "PCoA", distance = "unifrac")
weightedOrd <- ordinate(physeqStool, method = "PCoA", distance = "wunifrac")
```

```{r edit-samp-data}
# Change order of InfantMother (mother first)
sample_data(physeqStool)$InfantMother <- factor(sample_data(physeqStool)$InfantMother,
                                                levels = c("mother", "infant"))
sample_data(physeqStool)$InfantMother <- factor(sample_data(physeqStool)$InfantMother,
                                                labels = c("Mother", "Infant"))
```

```{r weighted-unifrac-pcoa, fig.width=9, fig.height=9}
weightedUniFracPCoA <- plot_ordination(physeqStool,
                                       ordination = weightedOrd,
                                       shape = "InfantMother",
                                       color = "InfantMother") +
  geom_point(size = 6) +
  scale_color_manual(values = c("#FFB8FF", "#50C367")) +
  scale_shape_manual(values = c(16, 15)) +
  geom_text_repel(aes_string(label = "Tarr_Subject_ID"),
                  size = 3.5, color = "gray30") +
  theme_bw() +
  theme(legend.position = "top",
        legend.text = element_text(size = 16),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_line(color = "black", size = 1.5),
        panel.border = element_rect(color = "black", fill = NA, size = 1.5))
  
weightedUniFracPCoA
```

```{r adonis}
set.seed(787)
RunAdonis(physeqStool, "InfantMother", "wunifrac")
```

# Pairwise Distance Boxplots Using Unweighted and Weighted UniFrac

Generate unweighted UniFrac pairwise distance data frame:

```{r unweightedDistDF}
set.seed(787)

# Calculate the distance object
unweightedDist <- UniFrac(physeqStool, weighted = FALSE)

# Coerce distance object into matrix
unweightedDistMtx <- as.matrix(unweightedDist)

# Melt and remove values of 0 to eliminate distances of each sample to itself
unweightedDistMtx_melt <- subset(reshape2::melt(unweightedDistMtx, 
                                                value.name = "UniFrac_Distance"),
                                 UniFrac_Distance != 0)

# Change Var1 and Var2 from factors to characters for easier use with dplyr
factorCols <- c("Var1", "Var2")
unweightedDistMtx_melt[factorCols] <- lapply(unweightedDistMtx_melt[factorCols], 
                                             as.character)

# Add columns for later subsetting when plotting.  This information will get
#   pulled from the metadata (metadataStool).  Within the *DistMtx_melt data 
#   frames, we need a column indicating whether Var1 and Var2 are infants,
#   mothers, and whether they're related.
unweightedDistDF <- unweightedDistMtx_melt

# Add Var1 and Var2 Type and Zygosity columns
# Put rownames back in metadataStool, then use the rownames to locate needed info
metadataStool <- column_to_rownames(metadataStool, var = "Sample")

unweightedDistDF <- unweightedDistDF %>%
  mutate(Var1_Type = metadataStool[unweightedDistDF$Var1, "InfantMother"],
         Var2_Type = metadataStool[unweightedDistDF$Var2, "InfantMother"])

# To easily determine the relatedness of the paired samples in each row,
#   create "short names" to replace the long sample names.
#   First put the rownames back into a column.

metadataStool <- rownames_to_column(metadataStool, var = "Sample")

lut_df <- dplyr::select(metadataStool, Sample, InfantMother, Tarr_Subject_ID) %>%
  mutate(uniqueID = paste(InfantMother, Tarr_Subject_ID, sep = ".")) %>%
  dplyr::select(Sample, uniqueID)

# convert lut_df into a vector
lut <- as.character(lut_df$uniqueID)
names(lut) <- lut_df$Sample

# Add in the short names to the distance data frames
unweightedDistDF$Var1_ShortName <- lut[unweightedDistDF$Var1]
unweightedDistDF$Var2_ShortName <- lut[unweightedDistDF$Var2]

# Add a column indicating whether a given row (pairwise comparison) is between
#   two related or unrelated samples.
#   Loop over the rows of the distance df, split Var1 and Var2 short names by 
#   "." delimiter and compare the 2nd element of the resulting list which will hold the 
#   Tarr Family ID,

AddPairRelation <- function(df) {
  
  oldDF <- df
  
  pairRelation <- vector(mode = "character", length = nrow(oldDF))
  
  for (i in 1:nrow(oldDF)) {
    Var1_split <- strsplit(x = df[i, "Var1_ShortName"], split = "\\.")
    Var2_split <- strsplit(x = df[i, "Var2_ShortName"], split = "\\.")
    
    if (Var1_split[[1]][[2]] == Var2_split[[1]][[2]]) {
      pairRelation[i] <- "related"
      } else {
        pairRelation[i] <- "unrelated"
      }
  }
  
  oldDF$Pair_Relation <- pairRelation
  
  return(oldDF)

}

unweightedDistDF <- AddPairRelation(unweightedDistDF)
```

Generate weighted UniFrac pairwise distance data frame:

```{r weightedDistDF}
# Repeat the same process for unweightedDistDF to create weightedDistDF
set.seed(787)

# Calculate weighted distance, melt and convert to df
weightedDist <- UniFrac(physeqStool, weighted = TRUE)

weightedDistMtx <- as.matrix(weightedDist)

weightedDistMtx_melt <- subset(reshape2::melt(weightedDistMtx,
                                              value.name = "WUniFrac_Distance"),
                               WUniFrac_Distance != 0)

weightedDistMtx_melt[factorCols] <- lapply(weightedDistMtx_melt[factorCols],
                                           as.character)

weightedDistDF <- weightedDistMtx_melt

# Add type and short names
metadataStool <- column_to_rownames(metadataStool, var = "Sample")

weightedDistDF <- weightedDistDF %>%
  mutate(Var1_Type = metadataStool[weightedDistDF$Var1, "InfantMother"],
         Var2_Type = metadataStool[weightedDistDF$Var2, "InfantMother"])

weightedDistDF$Var1_ShortName <- lut[weightedDistDF$Var1]
weightedDistDF$Var2_ShortName <- lut[weightedDistDF$Var2]

# Add pairwise relationships
weightedDistDF <- AddPairRelation(weightedDistDF)
```

## UniFrac Distances of Related and Unrelated Infants

```{r infantPairwiseBoxPlots, fig.width=9, fig.height=6}

PlotInfantDistances <- function(df, distanceVar, distanceLabel, statTest) {
  
  infantDF <- df %>%
    filter(Var1_Type == "infant" & Var2_Type == "infant")
  
  infantDFNoDupes <- infantDF[!duplicated(infantDF[distanceVar]), ]
  
  boxPlot <- MakeBoxPlot(df = infantDFNoDupes,
                         xVar = "Pair_Relation",
                         yVar = distanceVar,
                         label_y = distanceLabel,
                         statMethod = statTest)
}

infantPairwiseDistanceBoxPlots <- pmap(.l = list(df = list(unweighted = unweightedDistDF,
                                                           weighted = weightedDistDF),
                                                 distanceVar = list(unweighted = "UniFrac_Distance",
                                                                    weighted = "WUniFrac_Distance"),
                                                 distanceLabel = list(unweighted = "UniFrac Distance",
                                                                      weighted = "Weighted UniFrac Distance")),
                                       .f = PlotInfantDistances,
                                       statTest = "wilcox.test")
infantPairwiseDistanceBoxPlots
```

## UniFrac Distances of Related and Unrelated Infants and Mothers

```{r momInfPairwiseDistBoxPlots, fig.width=9, fig.height=6}
PlotMotherInfantDistances <- function(df, distanceVar, distanceLabel, statTest) {
  
  momAndInfDF <- df %>%
    filter(Var1_Type == "infant" & Var2_Type == "mother" |
           Var1_Type == "mother" & Var2_Type == "infant")
  
  momAndInfDFNoDupes <- momAndInfDF[!duplicated(momAndInfDF[distanceVar]), ]
  
  boxPlot <- MakeBoxPlot(df = momAndInfDFNoDupes,
                         xVar = "Pair_Relation",
                         yVar = distanceVar,
                         label_y = distanceLabel,
                         statMethod = statTest)
}

momInfPairwiseDistBoxPlots <- pmap(.l = list(df = list(unweighted = unweightedDistDF,
                                           weighted = weightedDistDF),
                                 distanceVar = list(unweighted = "UniFrac_Distance",
                                                    weighted = "WUniFrac_Distance"),
                                 distanceLabel = list(unweighted = "UniFrac Distance",
                                                      weighted = "Weighted UniFrac Distance")),
                       .f = PlotMotherInfantDistances,
                       statTest = "wilcox.test")
momInfPairwiseDistBoxPlots
```

```{r pairwise-dist-all-v-all}
# This plot needs to have three categories: all infants to all infants,
#   all mothers to all mothers, and all infants to all mothers.

# Remove duplicates from weightedDistDF, then subset.
#   Add a category column to indicate which group the subsets belong in then
#   bind back together for plotting.

# Remove Duplicates
weightedDistDFNoDupes <- weightedDistDF[!duplicated(weightedDistDF["WUniFrac_Distance"]), ]

# Add category columns 
weightedDistDFNoDupes <- weightedDistDFNoDupes %>%
  mutate(category = case_when(Var1_Type == "infant" & Var2_Type == "infant" ~ "Within Infants",
                              Var1_Type == "mother" & Var2_Type == "mother" ~ "Within Mothers",
                              TRUE ~ "Infants vs Mothers")) # if ELSE is TRUE?

```

```{r check-all-v-all-plot, fig.width=9, fig.height=6}
all_all_boxPlot <- MakeBoxPlot(df = weightedDistDFNoDupes,
                               xVar = "category",
                               yVar = "WUniFrac_Distance",
                               label_y = "Weighted UniFrac Distance",
                               statMethod = "kruskal.test")
all_all_boxPlot
```

```{r withinInfMoms, fig.width=9, fig.height=6}
# Check just the w/n infants vs w/n mothers
withinInfvMoms <- filter(weightedDistDFNoDupes,
                         category != "Infants vs Mothers")

withinInfvMomsPlot <- MakeBoxPlot(df = withinInfvMoms,
                                  xVar = "category",
                                  yVar = "WUniFrac_Distance",
                                  label_y = "Weighted UniFrac Distance",
                                  statMethod = "wilcox.test")
withinInfvMomsPlot
```

## Weighted UniFrac Distances between Mothers and Infants for C-section vs. Vaginal Deliveries (Related Pairs)

```{r dist-dlv-route, fig.width=9, fig.height=6}
# starting with weightedDistDFNoDupes
weightedDistRoute <- weightedDistDFNoDupes %>%
  filter(category == "Infants vs Mothers" & Pair_Relation == "related")

# In this case, Var1_Type is always infant and Var2_Type is always mother
#   so we can easily loop over the values for Var1 and look up the delivery route

metadataStool$DeliveryRoute <- as.character(metadataStool$DeliveryRoute)

weightedDistRoute <- weightedDistRoute %>%
  mutate("DeliveryRoute" = metadataStool[Var1, "DeliveryRoute"])

weightedPairwiseDistRoute <- MakeBoxPlot(df = weightedDistRoute,
                                         xVar = "DeliveryRoute",
                                         yVar = "WUniFrac_Distance",
                                         label_y = "Weighted UniFrac Distance",
                                         statMethod = "wilcox.test")
weightedPairwiseDistRoute

# Format this data frame for Prism
cSectionCol <- weightedDistRoute %>%
  filter(DeliveryRoute == "Cesarean") %>%
  dplyr::select(WUniFrac_Distance)
cSectionCol <- as.vector(cSectionCol$WUniFrac_Distance) # length = 12

vaginalCol <- weightedDistRoute %>%
  filter(DeliveryRoute == "Vaginal") %>%
  dplyr::select(WUniFrac_Distance)
vaginalCol <- as.vector(vaginalCol$WUniFrac_Distance) # length = 15

wUniFracDeliveryRoute = cbind("C-section" = c(cSectionCol,
                                              rep(NA,
                                                  length(vaginalCol) -
                                                    length(cSectionCol))),
                              "Vaginal" = cSectionCol)
```

## Weighted UniFrac Distances between Related Infants for Monozygotic vs. Dizygotic (related pairs)

```{r dist-zygosity, fig.width=9, fig.height=6}
weightedDistZygosity <- weightedDistDFNoDupes %>%
  filter(category == "Within Infants" & Pair_Relation == "related")

# Can simply used Var1 to look up the zygosity for the twin pair
metadataStool$Zygosity <- as.character(metadataStool$Zygosity)

weightedDistZygosity <- weightedDistZygosity %>%
  mutate("Zygosity" = metadataStool[Var1, "Zygosity"])

weightedPairwiseDistZygosity <- MakeBoxPlot(df = weightedDistZygosity,
                                            xVar = "Zygosity",
                                            yVar = "WUniFrac_Distance",
                                            label_y = "Weighted UniFrac Distance",
                                            statMethod = "wilcox.test")
weightedPairwiseDistZygosity

# Format this data frame for Prism
monozygoticColumn <- weightedDistZygosity %>%
  filter(Zygosity == "Monozygotic") %>%
  dplyr::select(WUniFrac_Distance)
monozygoticColumn <- as.vector(monozygoticColumn$WUniFrac_Distance) # length = 4

dizygoticColumn <- weightedDistZygosity %>%
  filter(Zygosity == "Dizygotic") %>%
  dplyr::select(WUniFrac_Distance)
dizygoticColumn <- as.vector(dizygoticColumn$WUniFrac_Distance) # length = 7

wUniFracZygosity = cbind("Monozygotic" = c(monozygoticColumn,
                                           rep(NA,
                                               length(dizygoticColumn) -
                                                 length(monozygoticColumn))),
                         "Dizygotic" = dizygoticColumn)
```

# Determine which samples belong to "Complete Families" - family groups having one mother and at least one infant

```{r find-complete-families}
# First determine which family groups have at least one mother and one infant sample.

metadataStool$InfantMother <- as.character(metadataStool$InfantMother)

allFamilyGroups <- as.character(unique(metadataStool$Family_Group))

keepFamilyGroupList <- list()

# loop over family groups and count how many moms and infants each group has
for (i in 1:length(allFamilyGroups)) {
  
  currentFamilyGroup <- allFamilyGroups[i]
  momCount <- 0
  infantCount <- 0
  keepFamilyGroup <- FALSE
  
  for (row in 1:nrow(metadataStool)) {
    if (metadataStool[row, "Family_Group"] == currentFamilyGroup) {
      if(metadataStool[row, "InfantMother"] == "mother") {
        momCount <- momCount + 1
        } else {
          infantCount <- infantCount + 1
      }
    }
    if (momCount > 0 & infantCount > 0) {
      keepFamilyGroup <- TRUE
    }
    keepFamilyGroupList[[currentFamilyGroup]] <- keepFamilyGroup
  }
}

keepFamilyGroupList <- unlist(keepFamilyGroupList)
keepFamilies <- names(keepFamilyGroupList[keepFamilyGroupList == TRUE])

# 10 families are incomplete.  Of these, families 11 and 13 are one infant, and
#   10 and 15 is two infants, removing 6 more infants from the data set for a 
#   total of 33 - 6 = 27

# Subset physeqStool

physeqCompleteFamilies <- subset_samples(physeqStool, 
                                         Family_Group %in% keepFamilies)

metadataCompleteFamilies <- data.frame(sample_data(physeqCompleteFamilies))
```

```{r save-metdataCompleteFamilies, eval=FALSE}
# Write out for future use
saveRDS(physeqCompleteFamilies, "../data/phyloseq_objects/physeqCompleteFamilies.RDS")

saveRDS(metadataCompleteFamilies, "../data/metadata/metadataCompleteFamilies.RDS")
```

# Circular Relative Abundance Barplot

```{r format-eyeball-plot}
phylumAbundance <- MakeAbundanceDF(physeqCompleteFamilies, 
                                   taxRank = "Phylum", 
                                   abundanceFilter = 0.0)

# For phyla with < 5% abundance, change to "other" category
phylumAbundance$Phylum <- as.character(phylumAbundance$Phylum) 
phylumAbdMut <- phylumAbundance %>%
  mutate("Name" = ifelse(test = Abundance < 0.05,
                         yes = "other", no = Phylum))
# Re-factor "Name"
phylumAbdMut$Name <- as.factor(phylumAbdMut$Name)
nameLevels <- levels(phylumAbdMut$Name)
nameLevels <- nameLevels[nameLevels != "other"]
nameLevels <- c(nameLevels, "other")

# Apply the new Name levels
phylumAbdMut$Name <- factor(phylumAbdMut$Name, levels = nameLevels)

# Assign colors: Set 2
myColors <- brewer.pal(n = length(nameLevels), name = "Set2")
names(myColors) <- nameLevels
```

```{r cutom-colors}
# Assign colors:
customColors <- brewer.pal(n = length(nameLevels), name = "Set1")
names(customColors) <- nameLevels
# add white for "none" category
customColors <- c(customColors, "none" = "#ffffff")

# Assign colors: cbPaletteGrey
greyPalette <- cbPaletteGrey[c(2,3,4,5,6,7,1)]
names(greyPalette) <- nameLevels
greyPalette <- c(greyPalette, "none" = "#ffffff")

# Assing colors: brewer
brewerPalette <- c('#8c510a','#bf812d','#dfc27d','#c7eae5','#80cdc1','#35978f','#01665e')
names(brewerPalette) <- nameLevels
brewerPalette <- c(brewerPalette, "non" = "#ffffff")
```

```{r eyeball-pieces-mother, fig.width=11, fig.height=8.5}
# Mothers - inner circle
phylumAbdMothers <- filter(phylumAbdMut, InfantMother == "mother")

# Order mothers and apply levels
motherLevels <- data.frame(sample_data(physeqCompleteFamilies))
motherLevels <- motherLevels %>% 
  rownames_to_column(var = "Sample") %>%
  filter(SampleType == "stool_mother") %>%
  dplyr::select(Sample, Tarr_Subject_ID)
motherLevels$Tarr_Subject_ID <- as.character(motherLevels$Tarr_Subject_ID)
motherLevels <- arrange(motherLevels, Tarr_Subject_ID)
motherLevelsVec <- as.factor(motherLevels$Sample)
phylumAbdMothers$Sample <- factor(phylumAbdMothers$Sample, levels = motherLevelsVec)

motherCircle <- ggplot(phylumAbdMothers,
                       aes(x = Sample, y = Abundance, fill = Name)) +
  geom_col()  +
  scale_fill_manual(name = "Name", values = customColors) +
  scale_y_continuous("", limits = c(-0.5, 1.1)) +
  coord_polar() +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank())+
  scale_x_discrete(labels = sampleLabelsVec)
motherCircle
```

```{r eyeball-pieces-infant, fig.width=11, fig.height=8.5}
#----- Infants -----#

phylumAbdInf <- filter(phylumAbdMut, InfantMother == "infant")

# Apply an ordering to the infants:
infantLevels <- data.frame(sample_data(physeqCompleteFamilies)) 
infantLevels <- infantLevels %>% 
  rownames_to_column(var = "Sample") %>%
  filter(SampleType == "stool_infant") %>%
  dplyr::select(Sample, Tarr_Subject_ID)
infantLevels$Tarr_Subject_ID <- as.character(infantLevels$Tarr_Subject_ID)
infantLevels <- arrange(infantLevels, Tarr_Subject_ID)
infantLevelsVec <- as.factor(infantLevels$Sample)
phylumAbdInf$Sample <- factor(phylumAbdInf$Sample, levels = infantLevelsVec)

infantBarPlot <- ggplot(phylumAbdInf,
                        aes_string(x = "Sample", 
                                   y = "Abundance", 
                                   fill = "Name")) +
  geom_col() +
  scale_fill_manual(name = "Name", values = myColors) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(labels = sampleLabelsVec)
infantBarPlot

# Add all infant samples as levels to this data frame so we get empty bars
#  where infants are missing (families with 1 or 2 infants)
# Make df with missing infants:
allInfants <- data.frame("Sample" = c("IC012.2", "IC019.1", "IC021.2", 
                                          "IC023.2", "IC025.2", "IC036.1", 
                                          "IC037.2", "IC045.1", "IC046.2"), 
                         "Tarr_Subject_ID" = c("C012.2", "C019.1", "C021.2", 
                                               "C023.2", "C025.2", "C036.1", 
                                               "C037.2", "C045.1", "C046.2"))

allInfants <- rbind(infantLevels, allInfants)
allInfants$Tarr_Subject_ID <- as.character(allInfants$Tarr_Subject_ID)
allInfants <- arrange(allInfants, Tarr_Subject_ID)
allInfants$Sample <- as.factor(allInfants$Sample)
# apply the "sample" columns as levels to phylumAbdInf
phylumAbdInf$Sample <- factor(phylumAbdInf$Sample,
                              levels = allInfants$Sample)
# Maybe add these samples to the data frame with 0's?
infPhyAbd <- phylumAbdInf %>% dplyr::select(Sample, Abundance, Name)
# Add the additional samples
missingInfants <- data.frame("Sample" = c("IC012.2", "IC019.1", "IC021.2", 
                                          "IC023.2", "IC025.2", "IC036.1", 
                                          "IC037.2", "IC045.1", "IC046.2"),
                             "Abundance" = 0,
                             "Name" = "none")
infPhyAbd <- rbind(infPhyAbd, missingInfants)

# Add "none" to myColors
myColorsPlusNone <- c(myColors, "none" = "#ffffff")

infantBarPlotAll <- ggplot(infPhyAbd,
                        aes_string(x = "Sample", 
                                   y = "Abundance", 
                                   fill = "Name")) +
  geom_col() +
  scale_fill_manual(name = "Name", values = myColorsPlusNone) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(labels = sampleLabelsVec)
infantBarPlotAll

# ciruclar plot
infantCircle <- ggplot(infPhyAbd,
                       aes(x = Sample, y = Abundance, fill = Name)) +
  geom_col() +
  scale_fill_manual(name = "Name", values = myColorsPlusNone) +
  scale_y_continuous("", limits = c(-1, 1.1)) +
  coord_polar() +
  theme_bw() +
  scale_x_discrete(labels = sampleLabelsVec) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank())
infantCircle

infantCircleCustomClr <- infantCircle + scale_fill_manual(name = "Name", 
                                                     values = customColors)
infantCircleCustomClr

```

# Run ADONIS on Metadata

Subset just the infants from physeqStool.

```{r physeqInfants}
physeqInfants <- physeqStool %>%
  subset_samples(SampleType == "stool_infant") %>%
  RemoveMissingTaxa()
physeqInfants # 1,470 taxa x 33 samples
```

```{r run-adonis}
variables <- c("DeliveryRoute", "FeedingType", "Zygosity",
               "DeliverySite", "PPBMICat", "DlvryBMICat",
               "Race")

set.seed(787)
map(variables, RunAdonis, distance = "wunifrac", physeq = physeqInfants)
```

***

### Session Info

```{r save-images}
imageName <- paste0("TwinMom16s_Analysis-", Sys.Date(), ".RData")
save.image(imageName)
```

```{r session-info}
Sys.Date()
getwd()
sessionInfo()
```

